Data Import, Clean, and Process

### 2025-03-13 notebooks/01-data-import-clean-process.qmd
reduced length. Moved content to appendices. Modified text. adjusted format. etc

set.seed(1234)

# load packages
library(tidyverse)

@@

```{r}
#| label: load-data

# load raw SPSS export (includes display order vars)
# downloaded on 2024-10-29 at 12:02 PM
raw_spss <- haven::read_sav("data-raw/vmf_election_worker_recruitment_survey_raw_spss_sanitized_20241029_T12.02.sav")

@@

A data dictionary contains metadata about the data. The `labelled::generate_dictionary` function is used to create a data dictionary (alternatively, use the `data_dict()` function from R package [surveytoolbox](https://martinctc.github.io/surveytoolbox/)). This function produces a dataframe into the R environment with the number of observations (rows) equal to number of variables (columns) in the original data set. For ease, I prefer to save the codebook/data dictionary as a `.csv` file. In order to do so, ensure that the function `labelled::convert_list_columns_to_character()` is run on the data dictionary which allows you to save the data dictionary as a `.csv` file. Alternatively, the
function `surveytoolbox::data_dict()` permits saving as a .csv file without the
additional step.

@@

# data dictionary from labelled package.
raw_spss_dict <- raw_spss |> 
  labelled::generate_dictionary() |> 
  # this allows me to save dictionary as .csv file
  labelled::convert_list_columns_to_character()

# alternative
# raw_spss_dict <- raw_spss |> surveytoolbox::data_dict()

# save data dictionary of raw_spss dataset download
write.csv(raw_spss_dict, file = "data/raw_spss_dict.csv")

@@

The data dictionary is the same as a codebook, but I also wanted to have a codebook that contained **only** the survey questions, variable labels (i.e., question text), responses (value labels), and associated values. So the saved `codebook` contains only the survey questions and the two grouping variables (`qset` and `group`).

@@

# this includes only the survey questions: question number (var), the question
# text (label_var), the value labels (label_val), and the numeric codes (value)
codebook <- raw_spss |> 
  janitor::clean_names() |> 
  # get rid of superfluous Qualtrics columns
  # Keep Qualtrics `response_id`
  select(!1:8 & !10:17 & !265) |>
  # subset spss data set to omit the display order variables
  select(!contains("_do_")
         & !contains("_ado_")
         & !contains("_click")
         & !contains("_count")) |>
  select(contains("q"), group) |>
  surveytoolbox::data_dict()

# save codebook as .csv file in data/ directory
write.csv(codebook, file = "data/survey_codebook.csv")

@@

From the raw data set, there are `r dim(raw_spss)[1]` observations in the data set (each observation is a row). The survey was programmed in such a way to automatically drop participants who,

@@

Although the survey was programmed in this way, however, any responses of these participants were still recorded and included in the data. Survey participants may not have been dropped from participating in the survey until after they had already entered in a few responses. For instance, a respondent failing both attention checks would not have been dropped from the survey until the tenth question (Q10) within the survey, but their responses prior to being dropped would still be included in a row within the data set. This results in a lot of missing data (i.e., NA values) filling up observations (rows) in the data. The following code filters the data set to exclude any respondent who met the aforementioned exclusion criteria.

>>>>

# set seed for reproducibility
set.seed(1234)

# load packages
library(tidyverse)
library(surveytoolbox)
library(labelled)
library(sjlabelled)
library(pewmethods)
library(fastDummies)

@@

Note that the 'raw' data set here has been sanitized of respondent's potentially sensitive information.

```{r}
#| label: load-data
#| echo: true

# load SPSS export from Qualtrics (includes display order vars)
# downloaded on 2024-10-29 at 12:02 PM
raw_spss_sanitized <- haven::read_sav("vmf_election_worker_recruitment_survey_raw_spss_sanitized_20241029_T12.02.sav")

@@

A data dictionary contains metadata about the data.
The the `data_dict()` function from R package [surveytoolbox](https://martinctc.github.io/surveytoolbox/) is used to create a *data dictionary* (i.e., codebook).
(alternatively, one can use the `generate_dictionary` from the [labelled](https://larmarange.github.io/labelled/) R package  to create a data dictionary with slightly different output.). This function produces a dataframe into the R environment with the number of observations (rows) equal to number of variables (columns) in the original data set. For ease, I prefer to save the codebook/data dictionary as a `.csv` file.

@@

# data dictionary from labelled package.
raw_spss_sanitized_data_dict <- raw_spss_sanitized |> 
  surveytoolbox::data_dict()

@@

The data dictionary is the same as a codebook, but I also wanted to have a codebook that contained **only** the survey questions, variable labels (`var`, i.e., question text), responses (`label_var`, i.e., value labels), and associated values (simply, `value`). So the saved `codebook` contains only the survey questions and the two grouping variables `qset` and `group`. The latter refers to the Experiment condition where `1 = Treatment` and `0 = Control`. The other grouping variable `qset` refers to two distinct question sets (A or B) which aren't as important to cover at the moment.

I run the code below to generate a codebook that contains only the survey question items (`var`), the question test (`label_var`), value labels (`label_val`) and corresponding numeric codes (`value`).

@@

# this includes only the survey questions: question number (var), the question
# text (label_var), the value labels (label_val), and the numeric codes (value)
codebook <- raw_spss_sanitized |> 
  janitor::clean_names() |> 
  # get rid of superfluous Qualtrics columns
  # Keep Qualtrics `response_id`
  select(!1:7, !9:11, -q_data_policy_violations) |>
  # subset spss data set to omit the display order variables
  select(!contains("_do_")
         & !contains("_ado_")
         & !contains("_click")
         & !contains("_count")) |>
  select(contains("q"), group) |>
  surveytoolbox::data_dict()

@@

From the raw data set, there are `r dim(raw_spss_sanitized)[1]` observations in the data set (each observation is a row). The survey was programmed in such a way to automatically drop participants who,

@@

Although the survey was programmed in this way, however, any responses of these participants were still recorded and included in the data. Survey participants may not have been dropped from participating in the survey until after they had already entered in a few responses, or at least responded to the first question requesting consent (Q1). Additionally, a respondent failing both attention checks would not have been dropped from the survey until the tenth question (Q10) within the survey, but their responses prior to being dropped would still be included in a row within the data set. This results in a lot of missing data (i.e., NA values) filling up observations (rows) in the data. The following code filters the data set to exclude any respondent who met the aforementioned exclusion criteria.
